Can We Formalise Type Theory Intrinsically without Any Compromise?

### 2025-12-03 tex/CPP26/final/main.tex
overfull hboxes

%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10003790.10011740</concept_id>
       <concept_desc>Theory of computation~Type theory</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

@@

\>[18]{}\mathbin{:}\;(\Varid{σ}\;\mathbin{:}\;\Conid{Sub}\;\Conid{Γ}\;\Conid{Δ})\;(\Varid{t}\;\mathbin{:}\;\Conid{Tm}\;\Conid{Γ})\;(\Varid{pt}\;\mathbin{:}\;\Varid{tyOf}\;\Varid{t}\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%

@@

@@

We extend our object type theory with dependent function types.
First we define the lifting of a substitution by a type as the following abbreviation:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{\char95 ↑\char95 }\;\mathbin{:}\;(\Varid{σ}\;\mathbin{:}\;\Conid{Sub}\;\Conid{Γ}\;\Conid{Δ})\;(\Conid{A}\;\mathbin{:}\;\Conid{Ty}\;\Conid{Δ})\;\Varid{→}\;\Conid{Sub}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}})\;(\Conid{Δ}\;\Varid{,}\;\Conid{A}){}\<[E]%
\\
\>[B]{}\Varid{\char95 ↑\char95 }\;\{\mskip1.5mu \Conid{Γ}\mskip1.5mu\}\;\Varid{σ}\;\Conid{A}\;\mathrel{=}\;{}\<[16]%
\>[16]{}\Varid{σ}\;\Varid{∘}\;\Varid{π₁}\;\{\mskip1.5mu \Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\mskip1.5mu\}\;\Varid{idS,}\;{}\<[E]%
\\
\>[16]{}\Varid{π₂}\;(\Varid{idS}\;\{\mskip1.5mu \Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\mskip1.5mu\})\;\Varid{∶[}\;\Varid{p}\;\mskip1.5mu]{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
where \ensuremath{\Varid{p}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{π₂}\;\Varid{idS})\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{∘}\;\Varid{π₁}\;\Varid{idS}\;\kern-1.5pt]_{\text{T}}}.
We can use \ensuremath{[\circ]_{\text{T}}} to define \ensuremath{\Varid{p}}, as \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;(\Varid{idS}\;\{\mskip1.5mu \Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\mskip1.5mu\}))} is equal to \ensuremath{\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{π₁}\;\Varid{idS}\;\kern-1.5pt]_{\text{T}}} by definition.\footnote{%
Yet, as interleaving function clauses with inductive types is also not supported, the strict equality \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;\Varid{σ})\;\mathrel{=}\;\Conid{A}\;[\kern-1.5pt\;\Varid{π₁}\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}} is not available at this point for \ensuremath{\Varid{p}}.
We again use forward declarations to introduce the required equalities \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;\Varid{σ})\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{π₁}\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}} and \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;\Varid{idS})\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{∘}\;\Varid{π₁}\;\Varid{idS}\;\kern-1.5pt]_{\text{T}}} to be defined later as \ensuremath{\Varid{refl}} and \ensuremath{[\circ]_{\text{T}}}, see \Cref{sec:tt:mutual} for further details.}

Other constructors are introduced following the `Ford transformation', with differences compared to the usual QIIT presentation highlighted:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{49}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\anonymous \;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Π}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Conid{A}\;\mathbin{:}\;\Conid{Ty}\;\Conid{Γ})\;(\Conid{B}\;\mathbin{:}\;\Conid{Ty}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}))\;\Varid{→}\;\Conid{Ty}\;\Conid{Γ}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\highlight{addition}{\Varid{app}}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Varid{t}\;\mathbin{:}\;\Conid{Tm}\;\Conid{Γ})\;(\Conid{B}\;\mathbin{:}\;\Conid{Ty}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}))\;\highlight{addition}{(\Varid{pt}\;\mathbin{:}\;\Varid{tyOf}\;\Varid{t}\;\mathop\equiv\;\Conid{Π}\;\Conid{A}\;\Conid{B})}\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Varid{→}\;\Conid{Tm}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{abs}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Varid{t}\;\mathbin{:}\;\Conid{Tm}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}))\;\Varid{→}\;\Conid{Tm}\;\Conid{Γ}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Π[]}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Conid{Π}\;\Conid{A}\;\Conid{B})\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\;\mathop\equiv\;\Conid{Π}\;(\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}})\;(\Conid{B}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑}\;\Conid{A}\;\kern-1.5pt]_{\text{T}}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\highlight{addition}{\Varid{abs[]}}\;{}\<[16]%
\>[16]{}\mathbin{:}\;\Varid{abs}\;\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}}\;\mathop\equiv\;\Varid{abs}\;(\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑}\;\Conid{A}\;\kern-1.5pt]_{\text{t}}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Πβ}\;{}\<[16]%
\>[16]{}\mathbin{:}\;\Varid{app}\;(\Varid{abs}\;\Varid{t})\;(\Varid{tyOf}\;\Varid{t})\;\highlight{addition}{\Varid{pt}}\;{}\<[49]%
\>[49]{}\mathop\equiv\;\Varid{t}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Πη}\;{}\<[16]%
\>[16]{}\mathbin{:}\;\Varid{abs}\;(\Varid{app}\;\Varid{t}\;\Conid{B}\;\Varid{pt})\;{}\<[49]%
\>[49]{}\mathop\equiv\;\Varid{t}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{tyOf}\;(\Varid{app}\;\Varid{t}\;\Conid{B}\;\Varid{p})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Conid{B}{}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Varid{abs}\;\{\mskip1.5mu \Conid{A}\;\mathrel{=}\;\Conid{A}\mskip1.5mu\}\;\Varid{t})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Conid{Π}\;\Conid{A}\;(\Varid{tyOf}\;\Varid{t}){}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Varid{abs[]}\;\Varid{σ}\;\Varid{t}\;\Varid{i})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Conid{Π[]}\;\Varid{σ}\;(\Varid{tyOf}\;\Varid{t})\;\Varid{i}{}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Conid{Πβ}\;\Varid{t}\;\Varid{pt}\;\Varid{i})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Varid{tyOf}\;\Varid{t}{}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Conid{Πη}\;\Varid{t}\;\Varid{pt}\;\Varid{i})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Varid{pt}\;(\mathord{\sim}\;\Varid{i}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

Finally, we introduce the equality constructor for the interaction between \ensuremath{\Varid{elim𝔹}} and substitution:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\anonymous \;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{elim𝔹[]}\;\mathbin{:}\;\Varid{...}\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{pt₂}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\mathop\equiv\;\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑𝔹}\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{tt}\;\Varid{∶[}\;[\mathsf{idS}]_{\text{T}}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{pu₂}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{u}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\mathop\equiv\;\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑𝔹}\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{ff}\;\Varid{∶[}\;[\mathsf{idS}]_{\text{T}}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{pb₂}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{b}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\mathop\equiv\;\Conid{𝔹}\;[\kern-1.5pt\;\Varid{idS}\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{q}\;\mathbin{:}\;{}\<[11]%
\>[11]{}\Conid{P}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{b}\;\Varid{∶[}\;\Varid{pb}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\;{}\<[E]%
\\
\>[11]{}\mathop\equiv\;\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{∘}\;\Varid{wk}\;\Varid{,}\;\Varid{vz}\;\Varid{∶[}\;\Conid{𝔹[]₂}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{b}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}}\;\Varid{∶[}\;\Varid{pb₂}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Varid{→}\;(\Varid{elim𝔹}\;{}\<[15]%
\>[15]{}\Conid{P}\;\Varid{t}\;\Varid{pt}\;\Varid{u}\;\Varid{pu}\;\Varid{b}\;\Varid{pb})\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}}\;{}\<[E]%
\\
\>[5]{}\hsindent{6}{}\<[11]%
\>[11]{}\mathop\equiv\;\Varid{elim𝔹}\;(\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑𝔹}\;\kern-1.5pt]_{\text{T}})\;{}\<[34]%
\>[34]{}(\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\Varid{pt₂}\;(\Varid{u}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\Varid{pu₂}\;{}\<[E]%
\\
\>[34]{}(\Varid{b}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\Varid{pb₂}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{tyOf}\;(\Varid{elim𝔹[]}\;\Conid{P}\;\Varid{u}\;\Varid{t}\;\Varid{pu}\;\Varid{pt}\;\Varid{b}\;\Varid{pb}\;\Varid{pt₂}\;\Varid{pu₂}\;\Varid{pb₂}\;\Varid{q}\;\Varid{i})\;\mathrel{=}\;\Varid{q}\;\Varid{i}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
Note again that we also defer the coherence proof of \ensuremath{\Varid{tyOf}} for \ensuremath{\Varid{elim𝔹[]}} by introducing another argument \ensuremath{\Varid{q}} in \ensuremath{\Varid{elim𝔹}} which can be removed when defining its elimination rule.

>>>>

%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
\begin{CCSXML}
   <concept>
       <concept_id>10003752.10003790.10011740</concept_id>
       <concept_desc>Theory of computation~Type theory</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

@@

\>[18]{}\mathbin{:}\;(\Varid{σ}\;\mathbin{:}\;\Conid{Sub}\;\Conid{Γ}\;\Conid{Δ})\;(\Varid{t}\;\mathbin{:}\;\Conid{Tm}\;\Conid{Γ})\;{}\<[E]%

@@

\>[18]{}\Varid{→}\;(\Varid{pt}\;\mathbin{:}\;\Varid{tyOf}\;\Varid{t}\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}})\;\Varid{→}\;\Conid{Sub}\;\Conid{Γ}\;(\Conid{Δ}\;\Varid{,}\;\Conid{A}){}\<[E]%

@@

We extend our object type theory with dependent function types.
First we define the lifting of a substitution by a type as the following abbreviation:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{6}{@{}>{\hspre}l<{\hspost}@{}}%
\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Varid{\char95 ↑\char95 }\;{}\<[6]%
\>[6]{}\mathbin{:}\;(\Varid{σ}\;\mathbin{:}\;\Conid{Sub}\;\Conid{Γ}\;\Conid{Δ})\;(\Conid{A}\;\mathbin{:}\;\Conid{Ty}\;\Conid{Δ})\;{}\<[E]%
\\
\>[6]{}\Varid{→}\;\Conid{Sub}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}})\;(\Conid{Δ}\;\Varid{,}\;\Conid{A}){}\<[E]%
\\
\>[B]{}\Varid{\char95 ↑\char95 }\;\{\mskip1.5mu \Conid{Γ}\mskip1.5mu\}\;\Varid{σ}\;\Conid{A}\;\mathrel{=}\;{}\<[16]%
\>[16]{}\Varid{σ}\;\Varid{∘}\;\Varid{π₁}\;\{\mskip1.5mu \Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\mskip1.5mu\}\;\Varid{idS,}\;{}\<[E]%
\\
\>[16]{}\Varid{π₂}\;(\Varid{idS}\;\{\mskip1.5mu \Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\mskip1.5mu\})\;\Varid{∶[}\;\Varid{p}\;\mskip1.5mu]{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
where \ensuremath{\Varid{p}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{π₂}\;\Varid{idS})\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{∘}\;\Varid{π₁}\;\Varid{idS}\;\kern-1.5pt]_{\text{T}}}.
We can use \ensuremath{[\circ]_{\text{T}}} to define \ensuremath{\Varid{p}}, as \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;(\Varid{idS}\;\{\mskip1.5mu \Conid{Γ}\;\Varid{,}\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\mskip1.5mu\}))} is equal to \ensuremath{\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{π₁}\;\Varid{idS}\;\kern-1.5pt]_{\text{T}}} by definition.\footnote{%
Yet, as interleaving function clauses with inductive types is also not supported, the strict equality \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;\Varid{σ})\;\mathrel{=}\;\Conid{A}\;[\kern-1.5pt\;\Varid{π₁}\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}} is not available at this point for \ensuremath{\Varid{p}}.
We again use forward declarations to introduce the required equalities \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;\Varid{σ})\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{π₁}\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}} and \ensuremath{\Varid{tyOf}\;(\Varid{π₂}\;\Varid{idS})\;\mathop\equiv\;\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{∘}\;\Varid{π₁}\;\Varid{idS}\;\kern-1.5pt]_{\text{T}}} to be defined later as \ensuremath{\Varid{refl}} and \ensuremath{[\circ]_{\text{T}}}, see \Cref{sec:tt:mutual} for further details.}

Other constructors are introduced following the `Ford transformation', with differences compared to the usual QIIT presentation highlighted:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{16}{@{}>{\hspre}l<{\hspost}@{}}%
\column{23}{@{}>{\hspre}l<{\hspost}@{}}%
\column{49}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\anonymous \;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Π}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Conid{A}\;\mathbin{:}\;\Conid{Ty}\;\Conid{Γ})\;(\Conid{B}\;\mathbin{:}\;\Conid{Ty}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}))\;\Varid{→}\;\Conid{Ty}\;\Conid{Γ}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\highlight{addition}{\Varid{app}}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Varid{t}\;\mathbin{:}\;\Conid{Tm}\;\Conid{Γ})\;(\Conid{B}\;\mathbin{:}\;\Conid{Ty}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}))\;{}\<[E]%
\\
\>[16]{}\Varid{→}\;\highlight{addition}{(\Varid{pt}\;\mathbin{:}\;\Varid{tyOf}\;\Varid{t}\;\mathop\equiv\;\Conid{Π}\;\Conid{A}\;\Conid{B})}\;\Varid{→}\;\Conid{Tm}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{abs}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Varid{t}\;\mathbin{:}\;\Conid{Tm}\;(\Conid{Γ}\;\Varid{,}\;\Conid{A}))\;\Varid{→}\;\Conid{Tm}\;\Conid{Γ}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Π[]}\;{}\<[16]%
\>[16]{}\mathbin{:}\;(\Conid{Π}\;\Conid{A}\;\Conid{B})\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\;\mathop\equiv\;\Conid{Π}\;(\Conid{A}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}})\;(\Conid{B}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑}\;\Conid{A}\;\kern-1.5pt]_{\text{T}}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\highlight{addition}{\Varid{abs[]}}\;{}\<[16]%
\>[16]{}\mathbin{:}\;\Varid{abs}\;\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}}\;\mathop\equiv\;\Varid{abs}\;(\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑}\;\Conid{A}\;\kern-1.5pt]_{\text{t}}){}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Πβ}\;{}\<[16]%
\>[16]{}\mathbin{:}\;\Varid{app}\;(\Varid{abs}\;\Varid{t})\;(\Varid{tyOf}\;\Varid{t})\;\highlight{addition}{\Varid{pt}}\;{}\<[49]%
\>[49]{}\mathop\equiv\;\Varid{t}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Conid{Πη}\;{}\<[16]%
\>[16]{}\mathbin{:}\;\Varid{abs}\;(\Varid{app}\;\Varid{t}\;\Conid{B}\;\Varid{pt})\;{}\<[49]%
\>[49]{}\mathop\equiv\;\Varid{t}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{tyOf}\;(\Varid{app}\;\Varid{t}\;\Conid{B}\;\Varid{p})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Conid{B}{}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Varid{abs}\;\{\mskip1.5mu \Conid{A}\;\mathrel{=}\;\Conid{A}\mskip1.5mu\}\;\Varid{t})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Conid{Π}\;\Conid{A}\;(\Varid{tyOf}\;\Varid{t}){}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Varid{abs[]}\;\Varid{σ}\;\Varid{t}\;\Varid{i})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Conid{Π[]}\;\Varid{σ}\;(\Varid{tyOf}\;\Varid{t})\;\Varid{i}{}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Conid{Πβ}\;\Varid{t}\;\Varid{pt}\;\Varid{i})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Varid{tyOf}\;\Varid{t}{}\<[E]%
\\
\>[B]{}\Varid{tyOf}\;(\Conid{Πη}\;\Varid{t}\;\Varid{pt}\;\Varid{i})\;{}\<[23]%
\>[23]{}\mathrel{=}\;\Varid{pt}\;(\mathord{\sim}\;\Varid{i}){}\<[E]%
\ColumnHook
\end{hscode}\resethooks

@@

Finally, we introduce the equality constructor for the interaction between \ensuremath{\Varid{elim𝔹}} and substitution:
\begin{hscode}\SaveRestoreHook
\column{B}{@{}>{\hspre}l<{\hspost}@{}}%
\column{3}{@{}>{\hspre}l<{\hspost}@{}}%
\column{5}{@{}>{\hspre}l<{\hspost}@{}}%
\column{11}{@{}>{\hspre}l<{\hspost}@{}}%
\column{15}{@{}>{\hspre}l<{\hspost}@{}}%
\column{34}{@{}>{\hspre}l<{\hspost}@{}}%
\column{E}{@{}>{\hspre}l<{\hspost}@{}}%
\>[B]{}\Keyword{data}\;\anonymous \;\Keyword{where}{}\<[E]%
\\
\>[B]{}\hsindent{3}{}\<[3]%
\>[3]{}\Varid{elim𝔹[]}\;\mathbin{:}\;\Varid{...}\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{pt₂}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\mathop\equiv\;\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑𝔹}\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{tt}\;\Varid{∶[}\;[\mathsf{idS}]_{\text{T}}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{pu₂}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{u}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\mathop\equiv\;\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑𝔹}\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{ff}\;\Varid{∶[}\;[\mathsf{idS}]_{\text{T}}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{pb₂}\;\mathbin{:}\;\Varid{tyOf}\;(\Varid{b}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\mathop\equiv\;\Conid{𝔹}\;[\kern-1.5pt\;\Varid{idS}\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}(\Varid{q}\;\mathbin{:}\;{}\<[11]%
\>[11]{}\Conid{P}\;[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{b}\;\Varid{∶[}\;\Varid{pb}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{T}}\;{}\<[E]%
\\
\>[11]{}\mathop\equiv\;\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{∘}\;\Varid{wk}\;\Varid{,}\;\Varid{vz}\;\Varid{∶[}\;\Conid{𝔹[]₂}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}}\;{}\<[E]%
\\
\>[11]{}\hsindent{4}{}\<[15]%
\>[15]{}[\kern-1.5pt\;\Varid{idS}\;\Varid{,}\;\Varid{b}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}}\;\Varid{∶[}\;\Varid{pb₂}\;\mskip1.5mu]\;\kern-1.5pt]_{\text{T}})\;{}\<[E]%
\\
\>[3]{}\hsindent{2}{}\<[5]%
\>[5]{}\Varid{→}\;(\Varid{elim𝔹}\;{}\<[15]%
\>[15]{}\Conid{P}\;\Varid{t}\;\Varid{pt}\;\Varid{u}\;\Varid{pu}\;\Varid{b}\;\Varid{pb})\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}}\;{}\<[E]%
\\
\>[5]{}\hsindent{6}{}\<[11]%
\>[11]{}\mathop\equiv\;\Varid{elim𝔹}\;(\Conid{P}\;[\kern-1.5pt\;\Varid{σ}\;\Varid{↑𝔹}\;\kern-1.5pt]_{\text{T}})\;{}\<[34]%
\>[34]{}(\Varid{t}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\Varid{pt₂}\;(\Varid{u}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\Varid{pu₂}\;{}\<[E]%
\\
\>[34]{}(\Varid{b}\;[\kern-1.5pt\;\Varid{σ}\;\kern-1.5pt]_{\text{t}})\;\Varid{pb₂}{}\<[E]%
\\[\blanklineskip]%
\>[B]{}\Varid{tyOf}\;(\Varid{elim𝔹[]}\;\Conid{P}\;\Varid{u}\;\Varid{t}\;\Varid{pu}\;\Varid{pt}\;\Varid{b}\;\Varid{pb}\;\Varid{pt₂}\;\Varid{pu₂}\;\Varid{pb₂}\;\Varid{q}\;\Varid{i})\;\mathrel{=}\;\Varid{q}\;\Varid{i}{}\<[E]%
\ColumnHook
\end{hscode}\resethooks
Note again that we also defer the coherence proof of \ensuremath{\Varid{tyOf}} for \ensuremath{\Varid{elim𝔹[]}} by introducing another argument \ensuremath{\Varid{q}} in \ensuremath{\Varid{elim𝔹}} which can be removed when defining its elimination rule.
